---
title: "CancerEcon Tutorial"
subtitle:  "Reading, Basic Indexing, and Merging Data"
author: Nick Fong (nick.fong [at])
date: 2020-06-25
output: html_document
---

## Library Imports
These need to be run to import other libraries, or collections of functions writen by other people to make our jobs easier
```{r}
suppressMessages(library(here))
library(readxl)
library(openxlsx)
library(tidyverse)
library(lubridate)
library(dplyr)
library(data.table)
```

## Read an Excel file and CSV file into separte dataframes `merged` and `country_codes`, respectively
```{r}
merged <- read_excel(here::here("allmerged.xlsx"), sheet = 1, guess_max = 5000)
country_codes <- read.csv(here::here("mergedcodes.csv"))
```
## Indexing and Assigning Variables:  Rename columns
Let's see what the names of the columns in our `country_codes` dataframe look like:
```{r}
colnames(country_codes)
```
We can rename the columns by indexing into the list of columns.  Each element in the list of columns has an index, starting with 1:
```{r}
colnames(country_codes)[1]
colnames(country_codes)[2]
```
What would you expect `colnames(country_codes)[3]` to output?
```{r}
colnames(country_codes)[3]
```
So we can rename the columns by assigning new names to the indexed elements:
```{r}
colnames(country_codes)[2] <- 'Name'
colnames(country_codes)[3] <- 'ISO3166'
```
We can also drop columns we don't care about in anticipation of our merge later.  The `%>%` is called a pipe; don't worry about what this is doing for now.  Note that the `select` function from `dplyr` chooses which columns you want to keep:
```{r}
country_codes <- country_codes %>% select(Name, ISO3166)
```
And now we can check our work:
```{r}
colnames(country_codes)
```

## Merge Data:  Add country names to merged data
Now we're going to add the country codes to our merged data so that it's easier for us to read.  To do this, we need a column in the `merged` dataframe that has the same name as the `country_code` data frame, so we'll start by renaming the relevant column in `merged`:
```{r}
colnames(merged)[2] <- 'ISO3166'
```
Next we should check to make sure that the column names are correct.  Verify that this is the case in the block below:
```{r}
# Your code here
```

Now we can merge the two data frames.  We're going to do a left outer join using the `merge` function.  This means that we want every row in the `merged` data frame to be preserved, and to ignore rows in the `country_codes` data frame that don't have a corresponding row in `merged`.  `Left` refers to the fact that we're preserving the rows in the left data frame, where left is given by the first argument to `merge` and right is given by the second argument.  If this doesn't make sense, don't worry about it for now.
```{r}
merged <- merge(merged, country_codes, by='ISO3166', all.x=TRUE)
```

It might be helpful to have the country names in the first column.  `dplyr` makes this easy for us:
```{r}
merged <- merged %>% select(Name, everything())
```

And now we can look at our data frame and see that we've added the country names to each row in our dataframe:
```{r}
merged
```
Finally, we can write our changes back to disk:
```{r}
write.csv(merged, file = here::here("merged_with_country_names.csv"), row.names = FALSE)
```
